---
title: "R Workbook 1"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

24 October, 2018

<br>

### Day 1

<br>

#### Introduction

The fundamentals of <b>R Programming Language</b> is discussed and illustrated in this Notebook.
Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

The chunk is denoted with notation starting with '''{r} and ending with ''' - The code will remain in between.

<br>

<b>Commonly used R data structures:</b>

1. Vectors - Contains elements of same type. c() function is often used to create vectors. Index starts at 1.
2. Lists - A generic vector containing elements of different type. Index starts at 1. Lists can be merged.
3. Matrices - Collection of data elements arranged in 2D rectangular layout. Contains element of same atomic type.
4. Arrays - Similar to matrices but can store data in ND. They can store data of same type only.
5. Factors - Used for fields that takes only pre-defined, finite number of values or categorical data. It store data as levels. It takes vectors as inputs.
6. Data Frames - Used for storing data tables. It is a list of vectors of equal length.

In the below section we will explore each one of them.

<b>Ex.1</b> Create a vector of 5 integers and fetch the 3rd, 4th and 5th element. Create another vector of 5 integers. Combine them to a new vector. Print the element wise sum and product of the two vectors. Create another vector of elements, each increases by a step of 0.2 i.e. 5.0 5.2 5.3 ... upto 7.0. Create a vector of strings of 4 Cities and sort them alphabetically. Give an example showing character vector as index. 

```{r}

# Creating vectors in two ways

vector.1 <- 1:5
vector.2 <- c(11:15)

vector.1
vector.2
typeof(vector.1)

# Combining two vectors
vector.3 <- c(vector.1, vector.2)
cat("Combined vector is:", vector.3, "\n")
typeof(vector.3)

# Vector arithmetic - Note that they must be of same length
length(vector.1)
vector.1+vector.2
vector.1*vector.2

# Create vector with sequence()
vector.4 <- seq(5, 7, by=0.2)
vector.4

# Vector of Strings and sorting
cities = c("London", "Paris", "Berlin", "Rome")
sort(cities)
sort(cities, decreasing = TRUE)

# Character vector as index - using names function
name.vector = c("Tom", "Lisa", "Harry")
name.vector
names(name.vector) = c("First", "Second", "Third")
names(name.vector)
# Retrieving the second member with its name
name.vector["Second"]

```

<b>Ex.2</b> Create a vector of 5 integers and fetch the 3rd, 4th and 5th element. Create another vector of 5 integers. Combine them to a new vector. Print the element wise sum and product of the two vectors. Create another vector of elements, each increases by a step of 0.2 i.e. 5.0 5.2 5.3 ... upto 7.0. Create a vector of strings of 4 Cities and sort them alphabetically. Give an example showing character vector as index.

```{r}

# Creating a List

list.1 = list(c(1, 2, 3), c("a", "b", "c", "d"), c(TRUE, FALSE, TRUE, TRUE))

# Accessing list elements
list.1[2]
list.1[c(1:3)]
list.1[[2]][2:4]

# Merging lists
list.2 = list(c("Ram", "Shyam", "Madhu"), c("Ganga", "Yamuna", "Narmada", "Sindhu"))
list.merged = c(list.1, list.2)
list.merged

```

<b>Examples on Matrix</b>

* Ex.3a - Create a matrix with 12 elements (may be 1-12 numbers) arranged by row
* Ex.3b - Create a matrix with 12 elements (may be 1-12 numbers) arranged by column
* Ex.3c - Create a matrix with 12 elements (may be 1-12 numbers) arranged in 4 rows and show the column and row names

```{r}

matrix.1 <- matrix(c(1:12), nrow = 4, byrow = TRUE)
matrix.1
matrix.2 <- matrix(c(1:12), nrow = 4, byrow = FALSE)
matrix.2

rnames = c("r1", "r2", "r3", "r4")
cnames = c("c1", "c2", "c3")

matrix.3 <- matrix(c(1:12), nrow = 4, byrow = TRUE, dimnames = list(rnames, cnames))
matrix.3

```

<b>Ex. 4</b> Retrieving / Accessing values from matrix.1:

* Get the element of 2nd row, 3rd column
* Get the entire 2nd row
* Get the entire 3rd column
* Get the 1st and 3rd columns
* Get the 3rd and 2nd rows
* Get the 1st & 2nd row and 2nd & 3rd column
* Get all rows except 1st
* Replace the value at 2R:2C with 100


```{r}

# We will operate on matrix.1 in above example

cat("The matrix:", "\n")
cat("--------------------------", "\n")
matrix.1
cat("--------------------------", "\n")

# Element of R2/C3
matrix.1[2,3]

# Entire R2
matrix.1[2,]

# Entire C3
matrix.1[,3]

# C1 & C3
matrix.1[,c(1,3)]

# R3 & R2
matrix.1[c(3,2),]

# R1 & R2 and C2 & C3
matrix.1[c(1,2), c(2,3)]

# All rows except R1
matrix.1[-1]

# Replace the value at 2R,2C with 100 and print
matrix.1[2,2] <- 100
matrix.1


```

