---
title: "R Cheatsheet 2"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

25 October, 2018

<br>

### Day 2

<br>

#### Matrices, Factors and Arrays

A matrix is a collection of data elements arranged in a 2-D rectangular layout. They contain elements of same atomic type. Syntax for creating a matrix: matrix(data, nrow, ncol, byrow, dimnames)

<br>

<b>Commonly used R data structures:</b>

1. Vectors - Contains elements of same type. c() function is often used to create vectors. Index starts at 1.
2. Lists - A generic vector containing elements of different type. Index starts at 1. Lists can be merged.
3. Matrices - Collection of data elements arranged in 2D rectangular layout. Contains element of same atomic type.
4. Arrays - Similar to matrices but can store data in ND. They can store data of same type only.
5. Factors - Used for fields that takes only pre-defined, finite number of values or categorical data. It store data as levels. It takes vectors as inputs.
6. Data Frames - Used for storing data tables. It is a list of vectors of equal length.

In the below section we will explore each one of them.

<b>Ex.1</b> Create a vector of 5 integers and fetch the 3rd, 4th and 5th element. Create another vector of 5 integers. Combine them to a new vector. Print the element wise sum and product of the two vectors. Create another vector of elements, each increases by a step of 0.2 i.e. 5.0 5.2 5.3 ... upto 7.0. Create a vector of strings of 4 Cities and sort them alphabetically. Give an example showing character vector as index. 

```{r}


```
<b>Ex.2</b> Create a vector of 5 integers and fetch the 3rd, 4th and 5th element. Create another vector of 5 integers. Combine them to a new vector. Print the element wise sum and product of the two vectors. Create another vector of elements, each increases by a step of 0.2 i.e. 5.0 5.2 5.3 ... upto 7.0. Create a vector of strings of 4 Cities and sort them alphabetically. Give an example showing character vector as index.

```{r}



```


<b>NOTE:</b> 

To be written.