---
title: "R Workbook 2"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

7 November, 2018

<br>

### Day 2

<br>

#### Miscellaneous Problems

<br>

R provides only very basics of string handling - if you want a very versatile and robust framework better look for Python. However, no computer language is complete without offering anything on string - so does R. Below examples will show some basic string manipulations (slicing and dicing) done using R. 
<br>

<b>NOTE</b> stringr is an important library in R. For more, read - https://rpubs.com/iPhuoc/stringr_manipulation

<b>Ex.1</b> Read the string 'CATACLISTM'and print the number of characters in it - use paste() function to concatenate Strings. Get another string from it as 'DOGACLISM'. Now, in the original string replace C's with D's i.e. convert the 'input string 'CATACLISM' into 'DATADLISM'. Count the number of vowels in the original string.

```{r}

str.1 = "CATACLISM"

# Doing substring replacement

print(sub("CAT", "DOG", str.1))

# Part 1: Note how we converted integer into String using toString() function 
# and then concatenated with a string with paste() function

print(paste("Number of characters in String",str.1,":", toString(nchar(str.1))))

# We need to replace C with D - use chartr() function
library(stringr)
str.2 <- chartr("C", "D", str.1)
print(str.2)

# Detecting the matches - using stringr
cat("-----------------------------------------------------", "\n")

# Initializing the counts
count.a = 0
count.e = 0
count.i = 0
count.o = 0
count.u = 0

if(str_detect(str.1, "a") || str_detect(str.1, "A")) {
  
  count.a = str_count(str.1, "A")+str_count(str.1,"a")
  cat("occurance of A/a:",count.a, "\n")  
}
if(str_detect(str.1, "e") || str_detect(str.1, "E")) {
  
  count.e = str_count(str.1, "E")+str_count(str.1,"e")
  cat("occurance of E/e:", count.e, "\n")  
}
if(str_detect(str.1, "i") || str_detect(str.1, "I")) {
  
  count.i = str_count(str.1, "I")+str_count(str.1,"i")
  cat("occurance of I/i:", count.i, "\n")  
}
if(str_detect(str.1, "o") || str_detect(str.1, "O")) {
  
  count.o = str_count(str.1, "O")+str_count(str.1,"o")
  cat("occurance of O/o:",count.o, "\n")  
}
if(str_detect(str.1, "u") || str_detect(str.1, "U")) {
  
  count.u = str_count(str.1, "U")+str_count(str.1,"u")
  cat("occurance of U/u:", count.u, "\n")  
}

cat("Total number of vowels in string", str.1, "=", count.a+count.e+count.i+count.o+count.u)

# Stripping / splitting and reversing the string - strsplit returns a list
str.splitted <- strsplit(str.1, NULL)[[1]]
str.splitted[1:length(str.splitted)]

# Getting the vector in reverse - concatenating and then printing
rev(str.splitted)
cat(paste(rev(str.splitted)))

```

<br>

<b>Factors</b>

Factors are used to represent categorical data - can be ordered or unordered. Factor => Integer vector where each has a label. Use factor() fuction to create factors. Factors are automatically created when we read a dataset using fuction like read.table() 

<b>Ex.1</b> Create a factor with some categorical data (say, yes/no or male/female). Show them in a table.

```{r}

response.fact <- factor(c("yes", "yes", "no", "yes", "no"))
response.fact

# Get it in a table - number of yes and no is displayed
table(response.fact)

```
<br>

<b>Missing Values - </b> Missing values are denoted by NA or NaN for undefined mathematical operations. Below functions are used to test such scenarios:

*is.na()* - Used to test objects if they are NA
*is.nan()* - Used to test for NaN
*NA values have a class also - integer NA, character NA etc*
*NaN is NA but reverse is not true*

```{r}

# Test the missing value

test.v <- c(1, 2, NA, 4, NaN, 6, 7)

# T for both NA and NaN
is.na(test.v)

# T for only NaN
is.nan(test.v)


```