---
title: "R Workbook 4"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

10 November, 2018

<br>

### Day 4

<br>

#### Data Visualization - Graphics

Following topics are demonstrated below:

*Barplot*
*Lineplot*
*Histogram*
*Pie Chart*
*Dot plot and Trendline*


<br>

<b>Bar Plot 1 - </b> Use function **barplot()**

```{r}

max.temp <- c(22, 24, 25, 25, 26, 24, 27)
# min.temp <- c(18, 18, 19, 18, 20, 22, 23)

# Below is just basic - keep it commented once understood - we will use the detail one
# barplot(max.temp)

# Provide a name to the chart and save the file as image
png("temp_chart.png")

# Below is the detail one - use it and keep this as uncommented
barplot(max.temp, main = "Max Temperature Variation Over a Week", xlab = "Days", ylab = "Max Temperature", names.arg = c("Mon", "Tue", "Wed", "Thu","Fri", "Sat", "Sun"), col = "#CEE95F", border = "red")

# Save the file
dev.off()

```

<br>
<b>Bar Plot 2 - </b>Print a barplot taking inputs from the student.csv file - student roll number vs. absent of class 6


```{r}

student.data <- read.csv("student.csv")

print(student.data)


# Getting the best attendance in class 5
best.attend.5 <- subset(student.data, days_absent == min(days_absent) & class == 5)
best.attend.5

absent.data <- student.data$days_absent
#barplot(absent.data.6)

absent.dataf.6 <- subset(student.data, days_absent & class == 6)
absent.dataf.6
barplot(absent.dataf.6$days_absent, main = "Absent Data for Class 6", xlab = "Roll Number", ylab = "Number of days absent", names.arg = c(absent.dataf.6$roll_number))


```

<br>
<b>Bar Plot 3 - </b>Print the above plot horizontally. Hint: use horiz = TRUE in parameter
```{r}

barplot(absent.dataf.6$days_absent, main = "Absent Data for Class 6", xlab = "Number of days absent", ylab = "Roll Number", names.arg = c(absent.dataf.6$roll_number), horiz = TRUE)

```

<br>
<b>Histogram Plot 4 - </b>Plot a histogram of wages of labourers - 400, 700, 200, 700, 400, 600, 500, 400, 500, 600, 700, 300, 600, 700, 400, 300). This will plot the wages of the workers and their frequency.

```{r}

wages = c(400, 700, 200, 700, 400, 600, 500, 400, 500, 600, 700, 300, 600, 700, 400, 300)

hist(wages, main = "Daily Wages", xlab = "Wages in INR", ylab = "Frequency", col = "pink", freq = TRUE)

```

<br>
<b>Histogram Plot 5 - </b>Plot the above histogram of wages of labourers but now with density and density line.

```{r}

wages = c(400, 700, 200, 700, 400, 600, 500, 400, 500, 600, 700, 300, 600, 700, 400, 300)

# Drawing the histogram
hist.d = hist(wages, main = "Daily Wages", xlab = "Wages in INR", ylab = "Density", xlim = c(200, 700), col = "yellow", freq = FALSE)

# Deawing the density line
lines(density(wages), col = "red")

```

<br>
<b>Line Plot 6 - </b>Plot a simple line graph with given x and y values.

```{r}

y = c(14, 12, 18, 10, 15)
x = c(1, 2, 3, 4, 5)

# Plot the Line chart
plot(x, y, type = "o", col = "red", xlab = "Month", ylab = "Rainfall in mm.", main = "Monthly Rainfall Chart")

```

<br>
<b>Line Plot 7 - </b>Plot the above rainfall chart, now for 3 different cities

```{r}

delhi = c(14, 12, 13, 10, 15)
kolkata = c(15, 10, 11, 15, 12)
mumbai = c(10, 13, 11, 12, 14)
data.month = c(1, 2, 3, 4, 5)

# Plot the line chart
plot(data.month, delhi, type = "o", col = "red", xlab = "Month", ylab = "Rainfall in mm.", main = "Monthly Rainfall Chart")
lines(data.month, kolkata, type = "o", col = "blue")
lines(data.month, mumbai, type = "o", col = "green")

legend("top", legend = c("Rainfall in Delhi", "Rainfall in Kolkata", "Rainfall in Mumbai"), col = 1:3, pch = 1)



```

<br>
<b>Pie Plot 8 - </b>Plot a pie chart showing monthly expenditure breakdown of a person e.g. housing 400, food 300, clothing 150, entertainment 100, others 200

```{r}

expenditure = c(400, 300, 150, 100, 200)

# Calculating the percentage - note how round() and sum() functions are used
pie.percent <- round(100*expenditure/sum(expenditure), 1)

# Adding % symbol to each p.c. value
pie.percent <- paste(pie.percent, "%", sep = "")

# Plotting the pie
pie(expenditure, labels = pie.percent, main = "Monthly Expenditure Breakdown", col = c("red", "orange", "yellow", "blue", "green"), border = "brown", clockwise = TRUE)

# Legend for the pie chart
legend("topright", c("Housing", "Food", "Clothing", "Entertainment", "Others"), cex = 0.8, fill = c("red", "orange", "yellow", "blue", "green"))




```

<br>
<b>Scatter Plot 9 - </b>Scatter plots are useful to find correlations among the data points. A best fit line, known as <b>trendline</b>, can be drawn in order to study the relationships.
<br>
<b>Problem:</b> use the dataset __mtcars__ available within R environment to create a basic scatterplot. we will use the columns __wt__ and __mpg__ for weight of the car and the miles per gallon (or milage).

```{r}

# Get the input values from the data set
input <- mtcars[,c('wt', 'mpg')]

plot(x = input$wt, y = input$mpg, xlab = "weight", ylab = "Milage", xlim = c(2.5, 5), ylim = c(15, 20), main = "Weight vs. Milage")

# Drawing the trendline - Note the use of Linear Model Function or lm()
abline(lm(input$mpg~ input$wt), col = "blue")

```

