---
title: "R Cheatsheet - Preliminaries"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

05 November, 2018

<br>

### Preliminary Exercises

__Ex.1__ Given any number n print the multiplication table with given m number of terms.
```{r}

tabnum = as.integer(readline(prompt = "Enter the number:"))
maxrange = as.integer(readline(prompt = "Enter the range:"))
tabcount = 1

while (tabcount<=maxrange) {
  
  product = tabnum*tabcount
  cat(tabnum,"X", tabcount, "=", product, "\n")
  tabcount = tabcount + 1
  
}

```

__Ex.2__ Given any number n (other than zero) check if it is even.

```{r}

mynum = as.integer(readline(prompt = "Enter the number:"))

if(mynum == 0) {
  print("Enter a non-zero integer!")
} else {
  
  if(mynum %% 2 == 0) {
     print("Even!")
  } else {
    print("Odd!")
  }
}

```

__Ex.3__ Given any three digit number n, find the unit, tens and hundreds place value.

```{r}

mynum.3= as.integer(readline(prompt = "Enter any 3-digit number:"))

mynum.hundreds = as.integer(mynum.3/100)
cat("Hundreds place value:", mynum.hundreds, "\n")

mynum.tensnum = mynum.3 - mynum.hundreds*100
mynum.tens = as.integer(mynum.tensnum/10)
cat("Tens place value:", mynum.tens, "\n")

mynum.ones = as.integer((mynum.tensnum - mynum.tens*10))
cat("Ones place value:", mynum.ones, "\n")

```


__Ex.4__ Given any number n print 10 subsequent numbers and finally print the sum.
<br>
Given, n = 1; print - 1 2 3 4 5 6 7 8 9 10 and <b>Sum</b> = 55

```{r}

# Use readline to read from console - it will be read as string
print(int <- readline(prompt = "Enter first number: "))

# Converting the string int to integer num
num <- as.integer(int)
count<-0
result<-0

while (count < 10) {
  
  cat("",num)
  result = result + num
  num = num + 1
  count = count + 1
  
}

cat("sum =", result)

```
__Ex.4a__ Given any number n, count and print in reverse direction upto 3,2,1. E.g. If n=10 print: 10 9 8 7 6 5 4 3 2 1

```{r}

num.start = as.integer(readline(prompt="Enter the start number: "))

while(num.start>0) {
  
  print(num.start)
  num.start = num.start - 1
}

```



__Ex.5__ Print the Fibonacci sequence upto 10 terms and print the value of the golden ratio.
Fibonacci Series is given as $F(n) = 0, 1, 1, 2, 3, 5, 8, 13, 21, 34, 55, ...$
<br>
where, t$_n$ = t$_{n-1}$ + t$_{n-2}$

```{r}

# take input from the user
nterms = as.integer(readline(prompt="How many terms? "))

# first two terms
n1 = 0
n2 = 1
count = 2

# check if the number of terms is valid
if(nterms <= 0) {
  print("Plese enter a positive integer")
} else {
  if(nterms == 1) {
    print("Fibonacci sequence:")
    print(n1)
  } else {
    print("Fibonacci sequence:")
    print(n1)
    print(n2)
    while(count < nterms) {
      nth = n1 + n2
      print(nth)
      
      # update values
      n1 = n2
      n2 = nth
      count = count + 1
    }
  }
}

```

