---
title: "R Workbook 2"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

7 November, 2018

<br>

### Day 3

<br>

#### Miscellaneous Problems

<br>

R provides only very basics of string handling - if you want a very versatile and robust framework better look for Python. However, no computer language is complete without offering anything on string - so does R. Below examples will show some basic string manipulations (slicing and dicing) done using R. 
<br>

<b>NOTE</b> stringr is an important library in R. For more, read - https://rpubs.com/iPhuoc/stringr_manipulation

<b>Ex.1</b> Read the string 'CATACLISTM'and print the number of characters in it - use paste() function to concatenate Strings. Get another string from it as 'DOGACLISM'. Now, in the original string replace C's with D's i.e. convert the 'input string 'CATACLISM' into 'DATADLISM'. Count the number of vowels in the original string.

```{r}

str.1 = "CATACLISM"

# Doing substring replacement

print(sub("CAT", "DOG", str.1))

# Part 1: Note how we converted integer into String using toString() function 
# and then concatenated with a string with paste() function

print(paste("Number of characters in String",str.1,":", toString(nchar(str.1))))

# We need to replace C with D - use chartr() function
library(stringr)
str.2 <- chartr("C", "D", str.1)
print(str.2)

# Detecting the matches - using stringr
cat("-----------------------------------------------------", "\n")

# Initializing the counts
count.a = 0
count.e = 0
count.i = 0
count.o = 0
count.u = 0

if(str_detect(str.1, "a") || str_detect(str.1, "A")) {
  
  count.a = str_count(str.1, "A")+str_count(str.1,"a")
  cat("occurance of A/a:",count.a, "\n")  
}
if(str_detect(str.1, "e") || str_detect(str.1, "E")) {
  
  count.e = str_count(str.1, "E")+str_count(str.1,"e")
  cat("occurance of E/e:", count.e, "\n")  
}
if(str_detect(str.1, "i") || str_detect(str.1, "I")) {
  
  count.i = str_count(str.1, "I")+str_count(str.1,"i")
  cat("occurance of I/i:", count.i, "\n")  
}
if(str_detect(str.1, "o") || str_detect(str.1, "O")) {
  
  count.o = str_count(str.1, "O")+str_count(str.1,"o")
  cat("occurance of O/o:",count.o, "\n")  
}
if(str_detect(str.1, "u") || str_detect(str.1, "U")) {
  
  count.u = str_count(str.1, "U")+str_count(str.1,"u")
  cat("occurance of U/u:", count.u, "\n")  
}

cat("Total number of vowels in string", str.1, "=", count.a+count.e+count.i+count.o+count.u)

# Stripping / splitting and reversing the string - strsplit returns a list
str.splitted <- strsplit(str.1, NULL)[[1]]
str.splitted[1:length(str.splitted)]

# Getting the vector in reverse - concatenating and then printing
rev(str.splitted)
cat(paste(rev(str.splitted)))

```

<br>
<b>Ex.2</b> Create a vector of 5 integers and fetch the 3rd, 4th and 5th element. Create another vector of 5 integers. Combine them to a new vector. Print the element wise sum and product of the two vectors. Create another vector of elements, each increases by a step of 0.2 i.e. 5.0 5.2 5.3 ... upto 7.0. Create a vector of strings of 4 Cities and sort them alphabetically. Give an example showing character vector as index.

```{r}



```


<b>NOTE:</b> 

To be written.