---
title: "R Workbook 3"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

10 November, 2018

<br>

### Day 3

<br>

#### Data Frames


<b>Data Frames - </b>Used to store tabular data in R. Package dplyr is very useful while working with data frames. Data frames are special types of list where every element of the list has to have the same length. Each element of the list => column and the length of each element of the list is the number of rows. Unlike matrices, data frames can store different classes of objects in each column. Data frames are usually created by reading in a dataset using the read.table() or read.csv() or they can be created explicitly with the data.frame() function. Data frames can be converted to a matrix by data.matrix(). 

```{r}

frame.1 <- data.frame(serial = 1:5, occupancy = c(T, T, F, F, T), gender = c("M", "F", "F", "F", "M"), city = c("Kolkata", "Mumbai", "Delhi", "Chennai", "Bengaluru" ), connectivity = c("air", "air", "air", "road", "rail"))

frame.1


```

<b>Exercise 1 - </b> Read data from a CSV file - student.csv. Print the highest score and show the full record for the student who got the highest score. Show the students of class 5 with best attendance ( meaning no. of days has to be minimum in collection)
```{r}

student.data <- read.csv("student.csv")

print(student.data)

# Getting the highest score i.e. a specific data
highest.score <- max(student.data$total_score)
print(highest.score)

# Getting the full record of highest score
ret.val <- subset(student.data, total_score == max(total_score))
print(ret.val)

# Getting the best attendance in class 5
best.attend.5 <- subset(student.data, days_absent == min(days_absent) & class == 5)
best.attend.5

```