---
title: "R Cheatsheet - Preliminaries"
output: html_notebook
---
Author: Avik Deb (avikdeb@gmail.com)

05 November, 2018

<br>

### Preliminary Exercises


__Ex.1__ Given any number n print 10 subsequent numbers and finally print the sum.
<br>
Given, n = 1; print - 1 2 3 4 5 6 7 8 9 10 and <b>Sum</b> = 55

```{r}

# Use readline to read from console - it will be read as string
print(int <- readline(prompt = "Enter first number: "))

# Converting the string int to integer num
num <- as.integer(int)
count<-0
result<-0

while (count < 10) {
  
  cat("",num)
  result = result + num
  num = num + 1
  count = count + 1
  
}

cat("sum =", result)

```
__Ex.2__ Print the Fibonacci sequence upto 10 terms and print the value of the golden ratio.
Fibonacci Series is given as $F(n) = 0, 1, 1, 2, 3, 5, 8, 13, 21, 34, 55, ...$
<br>
where, t$_n$ = t$_{n-1}$ + t$_{n-2}$

```{r}

# take input from the user
nterms = as.integer(readline(prompt="How many terms? "))

# first two terms
n1 = 0
n2 = 1
count = 2

# check if the number of terms is valid
if(nterms <= 0) {
  print("Plese enter a positive integer")
} else {
  if(nterms == 1) {
    print("Fibonacci sequence:")
    print(n1)
  } else {
    print("Fibonacci sequence:")
    print(n1)
    print(n2)
    while(count < nterms) {
      nth = n1 + n2
      print(nth)
      
      # update values
      n1 = n2
      n2 = nth
      count = count + 1
    }
  }
}

```


<b>NOTE:</b> 

Add a new chunk by clicking the *Insert R* button on the toolbar.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
